---
title: "IMW~RC"
author: "dan.auerbach@dfw.wa.gov"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    theme: flatly 
    code-fold: true
    toc: true
    toc-location: left
    grid:
      sidebar-width: 180px
      body-width: 1100px
      margin-width: 20px
encoding: UTF-8
---

next:
 - add drainage area and/or (relative distance from outlet?)
 - fit width models as a null comparison
 - other responses than pool_cv --> back to wood, thinking about site residuals and stream trt-ref differences in terms and coefs
 - fit by complex with stream as fixed or random effect? then compare to single stream fits
 - test interactions? more/fewer terms?

maybe:
 - rebuild hab measures from emap pull transects?
 - redo SSNblr from RC dgo objects?
 - spatial gams? https://noamross.github.io/gams-in-r-course/chapter3 

"integrating gradient and lateral constraint into basin-scale planning is no guarantee of success but it can help to reduce failure"

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
#| label: setup
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library("tidyverse", quietly = T)
library("sf")
library("mapview")
library("patchwork")
library("gt")

library("mgcv")
library("gratia")
library("marginaleffects")

theme_set(theme_minimal())

load(
  "~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_lwd_ssn/imw-rc.RData"
)
```

```{r envi, evaluate=FALSE}
#| label: build_envi

load(
  "~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_hab_251203.RData"
)

dir_data_common <- "~/T/DFW-Team WDFW Watershed Synthesis - data_common"
epsg <- 2927 #WA state standard; NAD83(HARN)/ft

sf_wrias <- sf::read_sf(file.path(dir_data_common, "WR_WAT_WRIA.gdb/")) |>
  mutate(
    wria = factor(WRIA_NR),
    wria_nr_nm = factor(paste0(
      str_pad(WRIA_NR, width = 2, pad = "0", ),
      "_",
      WRIA_NM
    ))
  ) |>
  select(WRIA_NR, WRIA_NM, wria, wria_nr_nm, acres = WRIA_AREA_ACR_QT) |>
  arrange(wria) |>
  sf::st_transform(sf::st_crs(epsg))

# HUC12 polys for IMW streams
# full set built in data_common/wsio/huc12getter.R via nhdplusTools::get_huc(AOI = sf::st_union(sf_wrias), type = 'huc12') |> select(huc12, name, areasqkm)
# here subset to IMW HUC12s with previously manually determined huc12 ids
sf_wa_huc12_imw <- readRDS(file.path(dir_data_common, "wbd/sf_wa_huc12.rds")) |>
  filter(
    huc12 %in%
      c(
        "171100210304", #Deep
        "171100210303", #W Twin
        "171100210302", #E Twin
        "171100180107", # BBC incl other HC
        "170800030605", #Mill
        "170800030604", #Abernathy
        "170800030603" #Germany
      )
  ) |>
  mutate(
    cmplx_desc = case_when(
      str_detect(name, "Deep|Twin") ~ "Straits of Juan de Fuca",
      str_detect(name, "Beef") ~ "Hood Canal",
      str_detect(name, "Mill|Aber|Ger") ~ "Lower Columbia"
    ),
    cmplx = case_when(
      cmplx_desc == "Straits of Juan de Fuca" ~ "ST",
      cmplx_desc == "Hood Canal" ~ "HC",
      cmplx_desc == "Lower Columbia" ~ "LC"
    )
  ) |>
  arrange(huc12) |>
  select(cmplx, cmplx_desc, huc12, huc12_nm = name, huc_area_sq_km = areasqkm)

# # NHDplus catchment polys, built in 'fifo_test_nwm.qmd' chunk 'sf_hr_h12_imw_cat'
# # intersected all cat centroids with IMW huc12 polys to get comids for IMW catchments
# # then further filtered to drop some 'marine' cats not in NWM
# # only needed if comids needed? for NWM or StreamCat or...?
# sf_cat_imw <- readRDS(
#   "~/T/DFW-Team WDFW Watershed Synthesis - General/fifo/sf_nhdphr_h12_imw_nhdp_cat.rds"
# )

# keeping things split by complex for now
sf_rc_rme_lc <- st_read(
  "~/T/DFW-Team WDFW Watershed Synthesis - data_common/rc/Metric_Engine-Metric_Engine_for_Germany_Creek-Frontal_Columbia_River/outputs/riverscapes_metrics.gpkg",
  layer = "vw_dgo_metrics"
) |>
  filter(huc12 %in% sf_wa_huc12_imw$huc12) |>
  rename(dgoid = dgoid.1) |>
  select(dgoid, everything(), -starts_with("dgoid.")) |>
  st_transform(epsg)

sf_rc_rme_st <- st_read(
  "~/T/DFW-Team WDFW Watershed Synthesis - data_common/rc/Metric_Engine-Metric_Engine_for_Salt_Creek-Frontal_Strait_of_Juan_De_Fuca/outputs/riverscapes_metrics.gpkg",
  layer = "vw_dgo_metrics"
) |>
  filter(huc12 %in% sf_wa_huc12_imw$huc12) |>
  rename(dgoid = dgoid.1) |>
  select(dgoid, everything(), -starts_with("dgoid.")) |>
  st_transform(epsg)

# combine IMW site measures and RC covars by complex
# manual inspection of sites not directly intersecting RC DGOs but distances are mostly very small
# then manual inspection of join with nearest feature to catch a few misassigned:
# ABR016 --> should get dgoid 4805, got 4674
# GER050 --> should get dgoid 14262, got 4179
# DEW539 --> should get dgoid 2338, got 1589
# doing two pass - intersect then nearest to reduce chance of joining sites to the wrong DGO metrics
# further inefficiency in the manual correction of a few misassigned sites that are near confluences or small tribs
# with rows_update failing for some reason so simply rebinding the corrected rows to the rest of the nearest join results
sf_site_rme_lc <- st_join(
  sf_site_meta |>
    filter(cmplx_abrv == "LC") |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_lc,
  join = st_intersects,
  largest = T
)

nearests <- st_join(
  sf_site_meta |>
    semi_join(
      as_tibble(sf_site_rme_lc) |> filter(is.na(dgoid)) |> distinct(site),
      by = "site"
    ) |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_lc,
  join = st_nearest_feature
)

nearests <- bind_rows(
  nearests |>
    filter(!(site %in% c("ABR016", "GER050"))),
  nearests |>
    filter(site %in% c("ABR016", "GER050")) |>
    mutate(
      dgoid = if_else(site == "ABR016", 4805, dgoid),
      dgoid = if_else(site == "GER050", 14262, dgoid)
    ) |>
    select(cmplx_strm:dgoid) |>
    left_join(
      as_tibble(sf_rc_rme_lc) |>
        filter(dgoid %in% c(4805, 14262)) |>
        select(-geom),
      by = "dgoid"
    )
)

# finally, combine the intersected and corrected nearest join results
sf_site_rme_lc <- bind_rows(
  drop_na(sf_site_rme_lc, dgoid),
  nearests
) |>
  arrange(cmplx_strm, site)

# now the Straits complex
sf_site_rme_st <- st_join(
  sf_site_meta |>
    filter(cmplx_abrv == "ST") |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_st,
  join = st_intersects,
  largest = T
)

nearests <- st_join(
  sf_site_meta |>
    semi_join(
      as_tibble(sf_site_rme_st) |> filter(is.na(dgoid)) |> distinct(site),
      by = "site"
    ) |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_st,
  join = st_nearest_feature
)

nearests <- bind_rows(
  nearests |>
    filter(!(site %in% c("DEW539"))),
  nearests |>
    filter(site %in% c("DEW539")) |>
    mutate(
      dgoid = if_else(site == "DEW539", 2338, dgoid)
    ) |>
    select(cmplx_strm:dgoid) |>
    left_join(
      as_tibble(sf_rc_rme_st) |>
        filter(dgoid == 2338) |>
        select(-geom),
      by = "dgoid"
    )
)

sf_site_rme_st <- bind_rows(
  drop_na(sf_site_rme_st, dgoid),
  nearests
) |>
  arrange(cmplx_strm, site)

rm(nearests)

# # final rechecks
# mapview(sf_rc_rme_lc, zcol = "prim_channel_gradient") +
#   mapview(sf_site_rme_lc, zcol = "prim_channel_gradient")

save.image(
  "~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_lwd_ssn/imw-rc.RData"
)
```

# fit by single stream 

## initial: combn formulas then devexpl/AIC rank

Begin by reproducing separate single basin fits of all covar combos, all smooths, no interactions, then compare deviance explained and AIC to examine best model(s) for each basin.

but need to dig more into why these fits might not be as good as the diagnostics suggest...

```{r}
lhs <- "pct_pools_cv"
rhs <- c(
  "prim_channel_gradient",
  "hect_vb_per_km",
  "low_lying_ratio",
  "q2",
  "sphigh"
)

mods <- tibble(
  form = unlist(
    #lapply(seq_along(v), function(i) combn(v, i, simplify = FALSE)),
    map(1:length(rhs), \(i) combn(rhs, i, simplify = F)),
    recursive = FALSE
  ) |>
    map(~ paste(lhs, "~ s(", paste(.x, collapse = ") + s("), ")"))
)
```

```{r fit_by_stream}
#| label: fit_by_stream
fits_by_stream <- bind_rows(
  as_tibble(sf_site_rme_lc) |>
    select(cmplx_strm, site, pct_pools_cv, all_of(rhs)), # |> summary()
  as_tibble(sf_site_rme_st) |>
    select(cmplx_strm, site, pct_pools_cv, all_of(rhs)) # |> summary()
) |>
  nest(.by = c(cmplx_strm)) |>
  cross_join(mods) |>
  mutate(
    fit = map2(
      form,
      data,
      ~ gam(
        as.formula(.x),
        data = .y,
        family = Gamma(link = "log"),
        method = "REML"
      )
    ),
    devexpl = map_dbl(fit, ~ summary(.x)[["dev.expl"]]),
    glance = map(fit, ~ broom::glance(.x))
  ) |>
  unnest(form) |>
  unnest(glance) |>
  arrange(cmplx_strm, desc(devexpl), AIC)
```

```{r fits_gt}
#| label: fits_gt
fits_by_stream |>
  select(cmplx_strm, form, devexpl, AIC, adj.r.squared) |>
  #slice_max(order_by = devexpl, n = 5, by = cmplx_strm) |>
  slice_min(order_by = AIC, n = 5, by = cmplx_strm) |>
  gt(groupname_col = "cmplx_strm") |>
  fmt_number(
    columns = c("devexpl", "AIC", "adj.r.squared"),
    decimals = 2
  ) |>
  opt_vertical_padding(scale = 0.5) |>
  opt_table_font(size = px(10))
```

```{r fits_appraise}
#| label: fits_appraise
g <- fits_by_stream |>
  filter(devexpl < 0.999) |> #arrange(desc(devexpl))
  slice_min(order_by = AIC, n = 1, by = cmplx_strm) |>
  select(cmplx_strm, form, devexpl, AIC, adj.r.squared, fit) |>
  mutate(
    check = pmap(
      list(fit, cmplx_strm, form, devexpl),
      \(m, s, frm, de) {
        gratia::appraise(m, method = "simulate") +
          plot_annotation(
            title = s,
            subtitle = paste(frm, "devexpl =", round(de, 2))
          )
      }
    )
  )

g$check
```

## revised: begin with all of interest and select = TRUE

and can better specify 'k' per smooth
and integrate 'local' covars (vlwd_cv, gravel)
and sets up better shift to fitting by complex with 'stream' as fixed or random effect (as both term and 'by=' in smooth)

```{r}
lhs <- "pct_pools_cv"
rhs <- c(
  #"vlwd_cv",
  "prim_channel_gradient",
  "hect_vb_per_km",
  "low_lying_ratio",
  "q2",
  "sphigh"
)
#as.formula(paste(lhs, "~ s(", paste(rhs, collapse = ") + s("), ")"))
# begin with default k for all smooths, then can adjust per smooth if needed
form <- paste(lhs, "~", paste0("s(", rhs, ")", collapse = " + "))

fits_by_stream <- bind_rows(
  as_tibble(sf_site_rme_lc),
  as_tibble(sf_site_rme_st)
) |>
  select(cmplx_strm, site, all_of(lhs), all_of(rhs)) |> # summary()
  nest(.by = c(cmplx_strm)) |>
  mutate(
    fit = map(
      data,
      ~ gam(
        as.formula(form),
        data = .x,
        family = Gamma(link = "log"),
        method = "REML",
        select = TRUE
      )
    ),
    devexpl = map_dbl(fit, ~ summary(.x)[["dev.expl"]]),
    glance = map(fit, ~ broom::glance(.x)),
    check = map(fit, ~ as_tibble(k.check(.x), rownames = "term"))
  ) |>
  unnest(glance) |>
  mutate(
    praise = pmap(
      list(fit, cmplx_strm, devexpl),
      \(mod, strm, dev) {
        gratia::appraise(mod, method = "simulate") +
          plot_annotation(
            title = strm,
            subtitle = paste(form, "\ndevexpl =", round(dev, 2))
          )
      }
    )
  )

fits_by_stream
```

Interesting or artifact that Germany the only stream with vlwd_cv not shrunk to 0?
Similarly, interesting that low_lying only retained in ETwin?
And q2 + sphigh only kept in Mill and Abe?
Note that larger edf values here do not necessarily correspond to 'signif' terms
see Abernathy for example, with gradient more 'sig' despite lower edf than hect_vb, and q2 'nonsig' despite edf close to 2

Could perhaps increase k for: 
 - gradient (maybe sphigh) in Deep
 - q2 in Mill
?

### diagnostics/effects

```{r kcheck_praise}
#| label: kcheck_praise

# fits_by_stream$check

# edf near zero, term was shrunk as uninformative
# across which streams?
fits_by_stream |>
  select(cmplx_strm, check) |>
  unnest(check) |>
  select(cmplx_strm, term, edf) |>
  pivot_wider(names_from = term, values_from = edf) |>
  gt(rowname_col = "cmplx_strm") |>
  tab_style_body(
    style = cell_fill(color = "orange"),
    columns = where(is.numeric), #starts_with("edf"),
    fn = \(x) x < 0.5
  ) |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  opt_vertical_padding(scale = 0.7) |>
  opt_table_font(size = px(11))

# pval low, maybe increase basis
fits_by_stream |>
  select(cmplx_strm, check) |>
  unnest(check) |>
  select(cmplx_strm, term, `p-value`) |>
  pivot_wider(names_from = term, values_from = c(`p-value`)) |>
  gt(rowname_col = "cmplx_strm") |>
  tab_style_body(
    style = cell_fill(color = "yellow"),
    columns = where(is.numeric), #starts_with("p-"),
    fn = \(x) x < 0.1
  ) |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  opt_vertical_padding(scale = 0.7) |>
  opt_table_font(size = px(11))

fits_by_stream$praise

map(fits_by_stream$fit, ~ summary(.x)) |> set_names(fits_by_stream$cmplx_strm)
```

second plot of the gradient-sphigh combined effect in Abernathy could be really nice if it holds?
```{r pred_plots}
#| label: pred_plots

plot_predictions(
  fits_by_stream$fit[[2]],
  #  condition = "prim_channel_gradient",
  #  condition = "hect_vb_per_km",
  condition = "sphigh",
  type = "response",
  points = 0.6,
  rug = TRUE
)

plot_predictions(
  fits_by_stream$fit[[2]],
  #  condition = c("prim_channel_gradient", "sphigh"),
  condition = c("sphigh", "prim_channel_gradient"),
  type = "response",
  points = 0.6,
  rug = TRUE
)

# partial plots for all signif smooths in each stream
fits_by_stream |>
  mutate(
    sigsmth = map(fit, ~ as_tibble(summary(.x)$s.table, rownames = "term"))
  ) |>
  select(cmplx_strm, fit, sigsmth) |>
  unnest(sigsmth) |>
  filter(`p-value` < 0.1) |>
  mutate(
    term = str_remove_all(term, "s\\(|\\)"),
    pp = pmap(
      list(fit, term, cmplx_strm, `p-value`),
      \(x, y, z, pv) {
        plot_predictions(
          x,
          condition = y,
          type = "response",
          points = 0.6,
          rug = TRUE
        ) +
          labs(subtitle = paste0(z, ", pval: ", round(pv, 4)))
      }
    )
  ) |>
  arrange(term) |>
  pull(pp) |>
  wrap_plots()
```

declare combo object 
```{r sf_dgopred}
#| label: sf_dgopred
sf_dgopred <- bind_rows(sf_rc_rme_lc, sf_rc_rme_st) |>
  left_join(
    select(as_tibble(sf_wa_huc12_imw), cmplx, huc12, huc12_nm) |>
      filter(cmplx != "HC") |>
      mutate(
        cmplx_strm = paste(cmplx, huc12_nm, sep = "_"),
        cmplx_strm = if_else(
          str_detect(cmplx_strm, "Mill|West"),
          paste0(cmplx_strm, "*"),
          cmplx_strm
        ) |>
          factor(
            levels = c(
              'HC_Stavis Creek*',
              'HC_Seabeck Creek',
              'HC_Big Beef Creek',
              'HC_Little Anderson Creek',
              'LC_Mill Creek*',
              'LC_Abernathy Creek',
              'LC_Germany Creek',
              'ST_Deep Creek',
              'ST_West Twin River*',
              'ST_East Twin River'
            )
          )
      ),
    by = "huc12"
  ) |>
  select(cmplx, huc12, cmplx_strm, all_of(rhs)) |> # summary()
  split(~cmplx_strm, drop = T) |>
  map(
    ~ .x |>
      mutate(
        est = predict(
          pull(filter(fits_by_stream, cmplx_strm == .x$cmplx_strm[1]), "fit")[[
            1
          ]],
          newdata = as_tibble(.x),
          type = "response"
        )
      ) |>
      drop_na(est)
  )
```

### support for prediction

site rme vals within stream distributions
reasonably nice, paneled by stream (better with log10 axes), and arguably better on bin2d
would be important to note that counts of dgo feature units are highest for smallest headwater/trib (low order) reaches
that were not the focus of sampling oriented to fish habitat
```{r}
# log10 axes are okay, only a single dgo in Deep Creek has gradient == 0
sf_dgopred |>
  # bind_rows() |>
  # as_tibble() |>
  # # mutate(across(c(prim_channel_gradient, hect_vb_per_km), log10)) |>
  # # summarise(sum(is.infinite(prim_channel_gradient)), sum(is.infinite(hect_vb_per_km)))
  # filter(prim_channel_gradient == 0)
  map(
    ~ .x |>
      ggplot(aes(prim_channel_gradient, hect_vb_per_km)) +
      #geom_point(size = 0.5, alpha = 0.2) +
      geom_bin2d(alpha = 0.6) +
      scale_fill_gradient(low = "grey90", high = "grey20") +
      geom_point(
        data = fits_by_stream |>
          filter(cmplx_strm == .x$cmplx_strm[1]) |>
          select(cmplx_strm, data) |>
          unnest(data),
        color = "blue",
        fill = "blue",
        shape = 17
      ) +
      scale_x_log10() +
      scale_y_log10() +
      labs(subtitle = .x$cmplx_strm[1])
  ) |>
  wrap_plots(nrow = 2)
```

scatter overlaid to see among stream overall overlap/diffs but also hard to see much?
```{r}
ggplot(
  mapping = aes(prim_channel_gradient, hect_vb_per_km, color = cmplx_strm)
) +
  geom_point(
    data = bind_rows(sf_dgopred) |>
      filter(prim_channel_gradient > 0, str_detect(cmplx_strm, "ST")),
    size = 0.5,
    alpha = 0.4
  ) +
  geom_point(
    data = fits_by_stream |>
      select(cmplx_strm, data) |>
      unnest(data) |>
      bind_rows() |>
      filter(str_detect(cmplx_strm, "ST")),
    shape = 17
  ) +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(values = pal_cmplx_strm, aesthetics = c("color", "fill"))
```

better as univariate densities with rug of site-reach
```{r}
# bind_rows(sf_dgopred) |>
#   filter(prim_channel_gradient > 0, str_detect(cmplx_strm, "ST")) |>
#   ggplot(aes(prim_channel_gradient, color = cmplx_strm, fill = cmplx_strm)) +
#   geom_density(alpha = 0.4, show.legend = F) +
#   geom_rug(
#     data = fits_by_stream |>
#       select(cmplx_strm, data) |>
#       unnest(data) |>
#       bind_rows() |>
#       filter(str_detect(cmplx_strm, "ST")),
#     sides = "b",
#     show.legend = F
#   ) +
#   scale_color_manual(values = pal_cmplx_strm, aesthetics = c("color", "fill")) +
#   facet_wrap(~cmplx_strm, ncol = 1)

sf_dgopred |>
  map(
    ~ .x |>
      ggplot(aes(
        prim_channel_gradient,
        color = cmplx_strm,
        fill = cmplx_strm
      )) +
      geom_density(alpha = 0.4, show.legend = F) +
      geom_rug(
        data = fits_by_stream |>
          filter(cmplx_strm == .x$cmplx_strm[1]) |>
          select(cmplx_strm, data) |>
          unnest(data),
        sides = "b",
        show.legend = F
      ) +
      scale_color_manual(
        values = pal_cmplx_strm,
        aesthetics = c("color", "fill")
      ) +
      labs(subtitle = .x$cmplx_strm[1])
  ) |>
  wrap_plots(nrow = 3, byrow = F)
```

### spatial/network prediction

need complex sf_dgo objects distinguished by stream/basin to apply per-stream fits to make prediction surfaces
could/should rebuild sf_rc_rme objects 
also cannot predict from models that include local/site covars
```{r}
map(sf_dgopred, summary)

sf_dgopred[1:3] |>
  map(
    ~ .x |>
      ggplot() +
      geom_sf(aes(fill = est, color = est)) +
      scale_color_viridis_b(
        option = "plasma",
        aesthetics = c("color", "fill"),
        breaks = quantile(.x$est, seq(0, 1, by = 0.25)),
        labels = round(quantile(.x$est, seq(0, 1, by = 0.25)), 3)
      ) +
      theme_void() +
      labs(
        title = "Predicted CV %pools",
        subtitle = paste(
          str_squish(paste0(
            format(
              pull(
                filter(fits_by_stream, cmplx_strm == .x$cmplx_strm[1]),
                "fit"
              )[[1]]$formula
            ),
            collapse = ""
          )),
          paste0(
            "Dev. expl.: ",
            scales::percent(
              pull(
                filter(fits_by_stream, cmplx_strm == .x$cmplx_strm[1]),
                "devexpl"
              ),
              2
            )
          ),
          sep = "\n"
        )
      )
  ) |>
  wrap_plots(nrow = 1)
```


## re-examine covar set

declare nonspatial combo metric object for all dgos
```{r}
rc <- bind_rows(sf_rc_rme_lc, sf_rc_rme_st) |>
  as_tibble() |>
  left_join(
    select(as_tibble(sf_wa_huc12_imw), cmplx, huc12, huc12_nm) |>
      filter(cmplx != "HC") |>
      mutate(
        cmplx_strm = paste(cmplx, huc12_nm, sep = "_"),
        cmplx_strm = if_else(
          str_detect(cmplx_strm, "Mill|West"),
          paste0(cmplx_strm, "*"),
          cmplx_strm
        ) |>
          factor(
            levels = c(
              # 'HC_Stavis Creek*',
              # 'HC_Seabeck Creek',
              # 'HC_Big Beef Creek',
              # 'HC_Little Anderson Creek',
              'LC_Mill Creek*',
              'LC_Abernathy Creek',
              'LC_Germany Creek',
              'ST_Deep Creek',
              'ST_West Twin River*',
              'ST_East Twin River'
            )
          )
      ),
    by = "huc12"
  ) |>
  drop_na(drainage_area)
```

RC RME values of drainage area some fairly strong but not >0.75 correlation with gradient and valley width (log10 transformed), but nearly perfectly with channel width (possible/likely that in fact channel width is calc'd/derived from drainage area via regional hydraulic geometry scaling?)
this does also informs the lognormal-ish distributions (at least of DA)
```{r}
rc |>
  select(
    cmplx_strm,
    drainage_area,
    prim_channel_gradient,
    hect_vb_per_km,
    channel_width
  ) |>
  #count(cmplx_strm) |> left_join(tibble(cmplx_strm = names(pal_cmplx_strm), val = pal_cmplx_strm))
  mutate(across(
    c(drainage_area, prim_channel_gradient, hect_vb_per_km, channel_width),
    log10
  )) |>
  filter(
    is.finite(drainage_area),
    is.finite(prim_channel_gradient),
    is.finite(hect_vb_per_km)
  ) |>
  GGally::ggpairs(mapping = aes(color = cmplx_strm, alpha = 0.15))
# scale_color_manual(
#   values = pal_cmplx_strm[5:10],
#   aesthetics = c("color", "fill")
# )
```

so still question of what including DA will do to other covars

# fit width_cv vs da + ...?

confirmatory null: quick first look says DA does most of the explaining for bankfull_width_mean as expected
though performance is somewhat uneven across streams
from very good in E Twin, Deep, Mill, to less amazing in W Twin, Abernathy
worth looking at residuals?
not (yet) pre-log10!

bankfull_cv also varied among streams and/but also consistently worse

vlwd_m2_cv devexpl and obs-vs-fit quick look sort of in between - maybe better across the board than bankfull_cv but worse than bankfull_mean

```{r}
#lhs <- "bankfull_width_mean"
#lhs <- "bankfull_cv"
lhs <- "vlwd_m2_cv"
rhs <- c(
  "drainage_area",
  "prim_channel_gradient",
  "hect_vb_per_km"
)

form <- paste(lhs, "~", paste0("s(", rhs, ")", collapse = " + "))

fits_by_stream_bankfull <- bind_rows(
  as_tibble(sf_site_rme_lc),
  as_tibble(sf_site_rme_st)
) |>
  mutate(
    bankfull_cv = bankfull_width_sd / bankfull_width_mean,
    vlwd_m2_cv = vlwd_per_m2_sd / vlwd_per_m2_mean
  ) |>
  select(cmplx_strm, site, all_of(lhs), all_of(rhs)) |> # summary()
  nest(.by = c(cmplx_strm)) |>
  mutate(
    fit = map(
      data,
      ~ gam(
        as.formula(form),
        data = .x,
        family = Gamma(link = "log"),
        method = "REML",
        select = TRUE
      )
    ),
    devexpl = map_dbl(fit, ~ summary(.x)[["dev.expl"]]),
    glance = map(fit, ~ broom::glance(.x)),
    check = map(fit, ~ as_tibble(k.check(.x), rownames = "term"))
  ) |>
  unnest(glance) |>
  mutate(
    praise = pmap(
      list(fit, cmplx_strm, devexpl),
      \(mod, strm, dev) {
        gratia::appraise(mod, method = "simulate") +
          plot_annotation(
            title = strm,
            subtitle = paste(form, "\ndevexpl =", round(dev, 2))
          )
      }
    )
  )

fits_by_stream_bankfull

fits_by_stream_bankfull$fit |>
  map(~ summary(.x)) |>
  set_names(fits_by_stream_bankfull$cmplx_strm)

fits_by_stream_bankfull$praise
```


# fit by complex

https://ecogambler.netlify.app/blog/interpreting-gams/#difference-smooths-how-does-a-smooth-effect-vary-among-groups

