---
title: "IMW~RC"
author: "dan.auerbach@dfw.wa.gov"
date: "`r Sys.Date()`"
format:
  html:
    embed-resources: true
    theme: flatly 
    code-fold: true
    toc: true
    toc-location: left
    grid:
      sidebar-width: 180px
      body-width: 1100px
      margin-width: 20px
encoding: UTF-8
---

next:
 - re-do fits by stream, stop doing all combos, do all covars (with select=T) and examine (then refit with retained?)
 - fit by complex with stream as fixed or random effect? then compare to single stream fits
 - add 'local' covars (e.g., vlwd_cv, pct_pools_mean, etc.) to see if they improve model fit
 - test interactions?

maybe:
 - rebuild hab measures from emap pull transects?
 - redo SSNblr from RC dgo objects?

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
#| label: setup
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library("tidyverse", quietly = T)
library("sf")
library("mapview")
library("patchwork")
library("gt")

library("mgcv")
library("gratia")
library("marginaleffects")

theme_set(theme_minimal())

load(
  "~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_lwd_ssn/imw-rc.RData"
)
```

```{r envi, evaluate=FALSE}
#| label: build_envi

load(
  "~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_hab_251203.RData"
)

dir_data_common <- "~/T/DFW-Team WDFW Watershed Synthesis - data_common"
epsg <- 2927 #WA state standard; NAD83(HARN)/ft

sf_wrias <- sf::read_sf(file.path(dir_data_common, "WR_WAT_WRIA.gdb/")) |>
  mutate(
    wria = factor(WRIA_NR),
    wria_nr_nm = factor(paste0(
      str_pad(WRIA_NR, width = 2, pad = "0", ),
      "_",
      WRIA_NM
    ))
  ) |>
  select(WRIA_NR, WRIA_NM, wria, wria_nr_nm, acres = WRIA_AREA_ACR_QT) |>
  arrange(wria) |>
  sf::st_transform(sf::st_crs(epsg))

# HUC12 polys for IMW streams
# full set built in data_common/wsio/huc12getter.R via nhdplusTools::get_huc(AOI = sf::st_union(sf_wrias), type = 'huc12') |> select(huc12, name, areasqkm)
# here subset to IMW HUC12s with previously manually determined huc12 ids
sf_wa_huc12_imw <- readRDS(file.path(dir_data_common, "wbd/sf_wa_huc12.rds")) |>
  filter(
    huc12 %in%
      c(
        "171100210304", #Deep
        "171100210303", #W Twin
        "171100210302", #E Twin
        "171100180107", # BBC incl other HC
        "170800030605", #Mill
        "170800030604", #Abernathy
        "170800030603" #Germany
      )
  ) |>
  mutate(
    cmplx_desc = case_when(
      str_detect(name, "Deep|Twin") ~ "Straits of Juan de Fuca",
      str_detect(name, "Beef") ~ "Hood Canal",
      str_detect(name, "Mill|Aber|Ger") ~ "Lower Columbia"
    ),
    cmplx = case_when(
      cmplx_desc == "Straits of Juan de Fuca" ~ "ST",
      cmplx_desc == "Hood Canal" ~ "HC",
      cmplx_desc == "Lower Columbia" ~ "LC"
    )
  ) |>
  arrange(huc12) |>
  select(cmplx, cmplx_desc, huc12, huc12_nm = name, huc_area_sq_km = areasqkm)

# # NHDplus catchment polys, built in 'fifo_test_nwm.qmd' chunk 'sf_hr_h12_imw_cat'
# # intersected all cat centroids with IMW huc12 polys to get comids for IMW catchments
# # then further filtered to drop some 'marine' cats not in NWM
# # only needed if comids needed? for NWM or StreamCat or...?
# sf_cat_imw <- readRDS(
#   "~/T/DFW-Team WDFW Watershed Synthesis - General/fifo/sf_nhdphr_h12_imw_nhdp_cat.rds"
# )

# keeping things split by complex for now
sf_rc_rme_lc <- st_read(
  "~/T/DFW-Team WDFW Watershed Synthesis - data_common/rc/Metric_Engine-Metric_Engine_for_Germany_Creek-Frontal_Columbia_River/outputs/riverscapes_metrics.gpkg",
  layer = "vw_dgo_metrics"
) |>
  filter(huc12 %in% sf_wa_huc12_imw$huc12) |>
  rename(dgoid = dgoid.1) |>
  select(dgoid, everything(), -starts_with("dgoid.")) |>
  st_transform(epsg)

sf_rc_rme_st <- st_read(
  "~/T/DFW-Team WDFW Watershed Synthesis - data_common/rc/Metric_Engine-Metric_Engine_for_Salt_Creek-Frontal_Strait_of_Juan_De_Fuca/outputs/riverscapes_metrics.gpkg",
  layer = "vw_dgo_metrics"
) |>
  filter(huc12 %in% sf_wa_huc12_imw$huc12) |>
  rename(dgoid = dgoid.1) |>
  select(dgoid, everything(), -starts_with("dgoid.")) |>
  st_transform(epsg)

# combine IMW site measures and RC covars by complex
# manual inspection of sites not directly intersecting RC DGOs but distances are mostly very small
# then manual inspection of join with nearest feature to catch a few misassigned:
# ABR016 --> should get dgoid 4805, got 4674
# GER050 --> should get dgoid 14262, got 4179
# DEW539 --> should get dgoid 2338, got 1589
# doing two pass - intersect then nearest to reduce chance of joining sites to the wrong DGO metrics
# further inefficiency in the manual correction of a few misassigned sites that are near confluences or small tribs
# with rows_update failing for some reason so simply rebinding the corrected rows to the rest of the nearest join results
sf_site_rme_lc <- st_join(
  sf_site_meta |>
    filter(cmplx_abrv == "LC") |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_lc,
  join = st_intersects,
  largest = T
)

nearests <- st_join(
  sf_site_meta |>
    semi_join(
      as_tibble(sf_site_rme_lc) |> filter(is.na(dgoid)) |> distinct(site),
      by = "site"
    ) |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_lc,
  join = st_nearest_feature
)

nearests <- bind_rows(
  nearests |>
    filter(!(site %in% c("ABR016", "GER050"))),
  nearests |>
    filter(site %in% c("ABR016", "GER050")) |>
    mutate(
      dgoid = if_else(site == "ABR016", 4805, dgoid),
      dgoid = if_else(site == "GER050", 14262, dgoid)
    ) |>
    select(cmplx_strm:dgoid) |>
    left_join(
      as_tibble(sf_rc_rme_lc) |>
        filter(dgoid %in% c(4805, 14262)) |>
        select(-geom),
      by = "dgoid"
    )
)

# finally, combine the intersected and corrected nearest join results
sf_site_rme_lc <- bind_rows(
  drop_na(sf_site_rme_lc, dgoid),
  nearests
) |>
  arrange(cmplx_strm, site)

# now the Straits complex
sf_site_rme_st <- st_join(
  sf_site_meta |>
    filter(cmplx_abrv == "ST") |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_st,
  join = st_intersects,
  largest = T
)

nearests <- st_join(
  sf_site_meta |>
    semi_join(
      as_tibble(sf_site_rme_st) |> filter(is.na(dgoid)) |> distinct(site),
      by = "site"
    ) |>
    select(
      cmplx_strm,
      site,
      starts_with("year"),
      ends_with("mean"),
      ends_with("sd"),
      -contains("per100"),
      -contains("vol"),
      -contains("W_D"),
      -contains("depth")
    ) |>
    mutate(
      pct_pools_cv = pct_pools_sd / pct_pools_mean,
      vlwd_cv = vlwd_sd / vlwd_mean
    ),
  sf_rc_rme_st,
  join = st_nearest_feature
)

nearests <- bind_rows(
  nearests |>
    filter(!(site %in% c("DEW539"))),
  nearests |>
    filter(site %in% c("DEW539")) |>
    mutate(
      dgoid = if_else(site == "DEW539", 2338, dgoid)
    ) |>
    select(cmplx_strm:dgoid) |>
    left_join(
      as_tibble(sf_rc_rme_st) |>
        filter(dgoid == 2338) |>
        select(-geom),
      by = "dgoid"
    )
)

sf_site_rme_st <- bind_rows(
  drop_na(sf_site_rme_st, dgoid),
  nearests
) |>
  arrange(cmplx_strm, site)

rm(nearests)

# # final rechecks
# mapview(sf_rc_rme_lc, zcol = "prim_channel_gradient") +
#   mapview(sf_site_rme_lc, zcol = "prim_channel_gradient")

save.image(
  "~/T/DFW-Team WDFW Watershed Synthesis - IMW analyses/imw_lwd_ssn/imw-rc.RData"
)
```

# fit by single stream 

## initial: combn formulas then devexpl/AIC rank

Begin by reproducing separate single basin fits of all covar combos, all smooths, no interactions, then compare deviance explained and AIC to examine best model(s) for each basin.

but need to dig more into why these fits might not be as good as the diagnostics suggest...

```{r}
lhs <- "pct_pools_cv"
rhs <- c(
  "prim_channel_gradient",
  "hect_vb_per_km",
  "low_lying_ratio",
  "q2",
  "sphigh"
)

mods <- tibble(
  form = unlist(
    #lapply(seq_along(v), function(i) combn(v, i, simplify = FALSE)),
    map(1:length(rhs), \(i) combn(rhs, i, simplify = F)),
    recursive = FALSE
  ) |>
    map(~ paste(lhs, "~ s(", paste(.x, collapse = ") + s("), ")"))
)
```

```{r fit_by_stream}
#| label: fit_by_stream
fits_by_stream <- bind_rows(
  as_tibble(sf_site_rme_lc) |>
    select(cmplx_strm, site, pct_pools_cv, all_of(rhs)), # |> summary()
  as_tibble(sf_site_rme_st) |>
    select(cmplx_strm, site, pct_pools_cv, all_of(rhs)) # |> summary()
) |>
  nest(.by = c(cmplx_strm)) |>
  cross_join(mods) |>
  mutate(
    fit = map2(
      form,
      data,
      ~ gam(
        as.formula(.x),
        data = .y,
        family = Gamma(link = "log"),
        method = "REML"
      )
    ),
    devexpl = map_dbl(fit, ~ summary(.x)[["dev.expl"]]),
    glance = map(fit, ~ broom::glance(.x))
  ) |>
  unnest(form) |>
  unnest(glance) |>
  arrange(cmplx_strm, desc(devexpl), AIC)
```

```{r fits_gt}
#| label: fits_gt
fits_by_stream |>
  select(cmplx_strm, form, devexpl, AIC, adj.r.squared) |>
  #slice_max(order_by = devexpl, n = 5, by = cmplx_strm) |>
  slice_min(order_by = AIC, n = 5, by = cmplx_strm) |>
  gt(groupname_col = "cmplx_strm") |>
  fmt_number(
    columns = c("devexpl", "AIC", "adj.r.squared"),
    decimals = 2
  ) |>
  opt_vertical_padding(scale = 0.5) |>
  opt_table_font(size = px(10))
```

```{r fits_appraise}
#| label: fits_appraise
g <- fits_by_stream |>
  filter(devexpl < 0.999) |> #arrange(desc(devexpl))
  slice_min(order_by = AIC, n = 1, by = cmplx_strm) |>
  select(cmplx_strm, form, devexpl, AIC, adj.r.squared, fit) |>
  mutate(
    check = pmap(
      list(fit, cmplx_strm, form, devexpl),
      \(m, s, frm, de) {
        gratia::appraise(m, method = "simulate") +
          plot_annotation(
            title = s,
            subtitle = paste(frm, "devexpl =", round(de, 2))
          )
      }
    )
  )

g$check
```

## revised: begin with all of interest and select = TRUE

and can better specify 'k' per smooth
and integrate 'local' covars (vlwd_cv, gravel)
and sets up better shift to fitting by complex with 'stream' as fixed or random effect (as both term and 'by=' in smooth)

```{r}
lhs <- "pct_pools_cv"
rhs <- c(
  "vlwd_cv",
  "prim_channel_gradient",
  "hect_vb_per_km",
  "low_lying_ratio",
  "q2",
  "sphigh"
)
#as.formula(paste(lhs, "~ s(", paste(rhs, collapse = ") + s("), ")"))
# begin with default k for all smooths, then can adjust per smooth if needed
form <- paste(lhs, "~", paste0("s(", rhs, ")", collapse = " + "))

fits_by_stream <- bind_rows(
  as_tibble(sf_site_rme_lc),
  as_tibble(sf_site_rme_st)
) |>
  select(cmplx_strm, site, all_of(lhs), all_of(rhs)) |> # summary()
  nest(.by = c(cmplx_strm)) |>
  mutate(
    fit = map(
      data,
      ~ gam(
        as.formula(form),
        data = .x,
        family = Gamma(link = "log"),
        method = "REML",
        select = TRUE
      )
    ),
    devexpl = map_dbl(fit, ~ summary(.x)[["dev.expl"]]),
    glance = map(fit, ~ broom::glance(.x)),
    check = map(fit, ~ as_tibble(k.check(.x), rownames = "term"))
  ) |>
  unnest(glance) |>
  mutate(
    praise = pmap(
      list(fit, cmplx_strm, devexpl),
      \(mod, strm, dev) {
        gratia::appraise(mod, method = "simulate") +
          plot_annotation(
            title = strm,
            subtitle = paste(form, "\ndevexpl =", round(dev, 2))
          )
      }
    )
  )

fits_by_stream
```

Interesting or artifact that Germany the only stream with vlwd_cv not shrunk to 0?
Similarly, interesting that low_lying only retained in ETwin?
And q2 + sphigh only kept in Mill and Abe?
Note that larger edf values here do not necessarily correspond to 'signif' terms
see Abernathy for example, with gradient more 'sig' despite lower edf than hect_vb, and q2 'nonsig' despite edf close to 2

Could perhaps increase k for: 
 - gradient (maybe sphigh) in Deep
 - q2 in Mill
?

```{r kcheck_praise}
#| label: kcheck_praise

# fits_by_stream$check

# edf near zero, term was shrunk as uninformative
# across which streams?
fits_by_stream |>
  select(cmplx_strm, check) |>
  unnest(check) |>
  select(cmplx_strm, term, edf) |>
  pivot_wider(names_from = term, values_from = edf) |>
  gt(rowname_col = "cmplx_strm") |>
  tab_style_body(
    style = cell_fill(color = "orange"),
    columns = where(is.numeric), #starts_with("edf"),
    fn = \(x) x < 0.5
  ) |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  opt_vertical_padding(scale = 0.7) |>
  opt_table_font(size = px(11))

# pval low, maybe increase basis
fits_by_stream |>
  select(cmplx_strm, check) |>
  unnest(check) |>
  select(cmplx_strm, term, `p-value`) |>
  pivot_wider(names_from = term, values_from = c(`p-value`)) |>
  gt(rowname_col = "cmplx_strm") |>
  tab_style_body(
    style = cell_fill(color = "yellow"),
    columns = where(is.numeric), #starts_with("p-"),
    fn = \(x) x < 0.1
  ) |>
  fmt_number(columns = where(is.numeric), decimals = 3) |>
  opt_vertical_padding(scale = 0.7) |>
  opt_table_font(size = px(11))

fits_by_stream$praise

map(fits_by_stream$fit, ~ summary(.x)) |> set_names(fits_by_stream$cmplx_strm)
```

second plot of the gradient-sphigh combined effect in Abernathy could be really nice if it holds?
```{r pred_plots}
#| label: pred_plots

summary(fits_by_stream$fit[[2]])

plot_predictions(
  fits_by_stream$fit[[2]],
  #  condition = "prim_channel_gradient",
  #  condition = "hect_vb_per_km",
  condition = "sphigh",
  type = "response",
  points = 0.6,
  rug = TRUE
)

plot_predictions(
  fits_by_stream$fit[[2]],
  #  condition = c("prim_channel_gradient", "sphigh"),
  condition = c("sphigh", "prim_channel_gradient"),
  type = "response",
  points = 0.6,
  rug = TRUE
)
```



# fit by complex

https://ecogambler.netlify.app/blog/interpreting-gams/#difference-smooths-how-does-a-smooth-effect-vary-among-groups

